Response to Pachter's Review

### 2020-01-29 index.Rmd
revision

---
title: "Response to Pachter’s Review"
author: "Joshua Paik and Igor Rivin"
date: '1/22/2020, Last Update: 1/28/2020 6:00 P.M. G.M.T.'
output:
  html_document:
    toc: yes
  pdf_document:
    keep_tex: yes
    toc: yes
---

>*"Academic politics are so vicious precisely because the stakes are so small."*
> - Henry Kissinger

@@

The following is copy and pasted from our paper (with edit in parentheses):

A permutation test is a nonparametric way of assessing the difference in mean between two populations. We are interested in whether an observed difference in mean is due to chance, and we can assess this in the following way.

0. Record the true difference in mean (or median) ($d\mu$).
1. $H_0: d\mu = 0, H_1: d\mu < 0$
2. Sample without replacement 1/2 of the combined data set (X) and what is left (Y)
3. Take the mean (or median) of X and Y and record the difference
4. Repeat 10,000 times and plot the histogram
5. Record the number of points (m) in the induced distribution that is more extreme than or equal to the observed $d\mu$. The probability m/10,000 is the probability that what was observed was due to chance.

```{r, echo = FALSE, include = FALSE}
# input data and the number of permutations
meanPermutation <- function(Data, n){
  output <- matrix(NA, ncol = 1, nrow = n)
  for(i in 1:n){
    #sample 1/2 of the data
    X_index <- sample(1:length(Data), floor(0.5 * length(Data)))
    Y_index <- setdiff(1:length(Data), X_index)
    X <- Data[X_index]
    Y <- Data[Y_index]
    #calculate the difference
    diff <- mean(X)-mean(Y)
    #store
    output[i, ] <- diff
  } 
  return(output)
}
```

```{r, echo = FALSE, include = FALSE}
# input data and the number of permutations
medianPermutation <- function(Data, n){
  output <- matrix(NA, ncol = 1, nrow = n)
  for(i in 1:n){
    #sample 1/2 of the data
    X_index <- sample(1:length(Data), floor(0.5 * length(Data)))
    Y_index <- setdiff(1:length(Data), X_index)
    X <- Data[X_index]
    Y <- Data[Y_index]
    #calculate the difference
    diff <- median(X)-median(Y)
    #store
    output[i, ] <- diff
  } 
  return(output)
}
```

@@

In the following figure, the vertical red bar is the observed difference for hypothesis 1 in the induced distribution, blue is the observed difference for hypothesis 2, and green is the observed difference for hypothesis 3.

```{r, echo=FALSE}
muA <- mean(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")))$amscit, na.rm = TRUE)
muB <- mean(filter(df, ((lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&(role=="professor")))$amscit, na.rm = TRUE)
muC <- mean(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor")))$amscit, na.rm = TRUE)

@@

set.seed(0)
dist <- meanPermutation(na.omit(df$amscit),10000)

hist(dist,
     main = "Permutation Test on MSN cits only R1 Math Professors",
     xlab = "Differences in Mean")
abline(v=val1, col = "red")
abline(v=val2, col = "blue")
abline(v=val3, col = "green")

```
```{r, echo=FALSE, include = FALSE}
sum(dist < val1)/length(dist)
sum(dist < val2)/length(dist)
sum(dist < val3)/length(dist)
```

The induced p-value for hypothesis 1 is 0. The induced p-value for hypothesis 2 is 0.0002. The induced p-value for hypothesis 3 is 0. Hence we reject all three null hypotheses in favor of the alternative, and $\mu(A) < \mu(B) < \mu(C)$.

@@

As above, the vertical red bar is the observed difference for hypothesis 1 in the induced distribution, blue is the observed difference for hypothesis 2, and green is the observed difference for hypothesis 3.

```{r, echo = FALSE}
muA <- mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$amscitperyear, na.rm = TRUE)
muB <- mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$amscitperyear, na.rm = TRUE)
muC <- mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$amscitperyear, na.rm = TRUE)

@@

set.seed(0)
dist <- meanPermutation(na.omit(df$amscitperyear),10000)

hist(dist,
     main = "Permutation Test MSN citperyear R1 Math Professors",
     xlab = "Differences in Mean")
abline(v=val1, col = "red")
abline(v=val2, col = "blue")
abline(v=val3, col = "green")
```
```{r, echo=FALSE, include = FALSE}
sum(dist < val1)/length(dist)
sum(dist < val2)/length(dist)
sum(dist < val3)/length(dist)
```

The induced p-value for hypothesis 1 is 0. The induced p-value for hypothesis 2 is 0.0099. The induced p-value for hypothesis 3 is 0. Hence we reject all three null hypotheses, even assessing the induced p-value at a 2% significance level, in favor of the alternative, and we conclude that $\mu(A_{citperyear}) < \mu(B_{citperyear}) < \mu(C_{citperyear})$.

One may object of our usage of the mean here, opposed to the median. So we reperform the permutation test with the median. All three induced p-values are 0 so we can reject all three null hypotheses when comparing median citations per year.

```{r, echo=FALSE, include = FALSE}
muA <- median(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$amscitperyear, na.rm = TRUE)
muB <- median(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$amscitperyear, na.rm = TRUE)
muC <- median(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$amscitperyear, na.rm = TRUE)

set.seed(0)
dist <- medianPermutation(na.omit(df$amscitperyear),10000)
```

```{r, echo=FALSE, include = FALSE}
sum(dist < val1)/length(dist)
sum(dist < val2)/length(dist)
sum(dist < val3)/length(dist)
```

@@

As above, the vertical red bar is the observed difference for hypothesis 1 in the induced distribution, blue is the observed difference for hypothesis 2, and green is the observed difference for hypothesis 3.

```{r, echo = FALSE}
muA <- mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$amscitperyearpower, na.rm = TRUE)
muB <- mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$amscitperyearpower, na.rm = TRUE)
muC <- mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$amscitperyearpower, na.rm = TRUE)

@@

set.seed(0)
dist <- meanPermutation(na.omit(df$amscitperyearpower),10000)

hist(dist,
     main = "Permutation Test MSN citperyear^1.3 R1 Math Professors",
     xlab = "Differences in Mean")
abline(v=val1, col = "red")
abline(v=val2, col = "blue")
abline(v=val3, col = "green")
```
```{r, echo=FALSE, include = FALSE}
sum(dist < val1)/length(dist)
sum(dist < val2)/length(dist)
sum(dist < val3)/length(dist)
```

The induced p-value for hypothesis 1 is 0. The induced p-value for hypothesis 2 is 0.0207. The induced p-value for hypothesis 3 is 0. Hence we fail to reject hypothesis 2 at a 2% significance level and reject hypotheses 1 and 3 in favor of the alternative. We conclude that after adjusting for age $\mu(A) < \mu(B) \leq \mu(C)$.

@@

```{r, echo = FALSE, include = FALSE}
mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$amscitperyear, na.rm = TRUE)
mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$amscitperyear, na.rm = TRUE)
mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$amscitperyear, na.rm = TRUE)
```

@@

2. For letter B - 41.12345

>>>>

---
title: "Response to Pachter’s Review"
author: "Joshua Paik and Igor Rivin"
date: '1/22/2020, Last Update: 1/28/2020 6:00 P.M. G.M.T.'
output:
  pdf_document:
    keep_tex: yes
    toc: yes
  html_document:
    toc: yes
---

@@

A permutation test is a non-parametric means of assessing the significance test of a population. Throughout this section, we will be comparing the mean citations of two populations, X and Y. We will work under the assumption that our null hypothesis is $H_0: \mu(X) = \mu(Y)$, and our alternative is $H_1: \mu(X) < \mu(Y)$.

A permutation test works as follows. Let $X$ and $Y$ be our relevant populations, of size $n_X$ and $n_Y$. We would like to know whether we can accept that the observed difference in means was not due to chance. We record the observed difference in means as $\delta = \mu(X) - \mu(Y)$. We then take the union of our two population, $Z = X \cup Y$, and randomly partition $Z$ into two new sets $A$ and $B$, where $|A| = n_X$ and $|B| = n_Y$. We store $\mu(X)-\mu(Y)$ and induce a distribution $D$ of potential differences and repeat the process $n=10,000$ times. We can induce the p-value, or the probability that our observed difference was due to chance by the probability $p = |\{d:d\leq \delta, \forall d \in D\}|/n$.

```{r, echo = FALSE}
# input data and the number of permutations
meanPermutation <- function(Data, n,n_a){
  output <- matrix(NA, ncol = 1, nrow = n)
  p<-n_a/length(Data)
  for(i in 1:n){
    #sample p% of the data
    X_index <- sample(1:length(Data), floor(p * length(Data)))
    Y_index <- setdiff(1:length(Data), X_index)
    X <- Data[X_index]
    Y <- Data[Y_index]
    #calculate the difference
    diff <- mean(X)-mean(Y)
    #store
    output[i, ] <- diff
  } 
  return(output)
}
```

```{r, echo = FALSE}
# input data and the number of permutations
medianPermutation <- function(Data, n,n_a){
  output <- matrix(NA, ncol = 1, nrow = n)
  p<-n_a/length(Data)
  for(i in 1:n){
    #sample p% of the data
    X_index <- sample(1:length(Data), floor(p * length(Data)))
    Y_index <- setdiff(1:length(Data), X_index)
    X <- Data[X_index]
    Y <- Data[Y_index]
    #calculate the difference
    diff <- median(X)-median(Y)
    #store
    output[i, ] <- diff
  } 
  return(output)
}
```

@@

```{r, include = FALSE, echo = FALSE}
muA <- mean(filter(df, ((lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor")))$amscit, na.rm = TRUE)
muB <- mean(filter(df, ((lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&(role=="professor")))$amscit, na.rm = TRUE)
muC <- mean(filter(df, ((lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor")))$amscit, na.rm = TRUE)

@@

set.seed(0)
pop <- filter(df, (institution=="domesticr1"&field=="math"&role=="professor"))

n_a <- length(na.omit(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&(role=="professor"))$amscit))
n_b <- length(na.omit(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&(role=="professor"))$amscit))
n_c <- length(na.omit(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&(role=="professor"))$amscit))

dist1 <- meanPermutation(na.omit(filter(df,(lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "B and C"|lettergroup == "B Only"))$amscit),10000,n_a)
dist2 <- meanPermutation(na.omit(filter(df,lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "C Only")$amscit),10000,n_b)
dist3 <- meanPermutation(na.omit(filter(df,lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "C Only"|lettergroup == "B and C")$amscit),10000,n_c)

sum(dist1 < val1)/length(dist1)
sum(dist2 < val2)/length(dist2)
sum(dist3 < val3)/length(dist3)
```

The induced p-value for hypothesis 1 is 0. The induced p-value for hypothesis 2 is 0.0016. The induced p-value for hypothesis 3 is 0. Hence we reject all three null hypotheses in favor of the alternative, and $\mu(A) < \mu(B) < \mu(C)$.

@@

```{r, include = FALSE, echo = FALSE}
muA <- mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE)
muB <- mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE)
muC <- mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear, na.rm = TRUE)

@@

set.seed(0)

n_a <- length(na.omit(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear))
n_b <- length(na.omit(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear))
n_c <- length(na.omit(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyear))

pop<- filter(df, ((role=="professor")&field=="math"&institution=="domesticr1"))

dist1 <- meanPermutation(na.omit(filter(pop,(lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "B and C"|lettergroup == "B Only"))$amscitperyear),10000,n_a)
dist2 <- meanPermutation(na.omit(filter(pop,lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "C Only")$amscitperyear),10000,n_b)
dist3 <- meanPermutation(na.omit(filter(pop,lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "C Only"|lettergroup == "B and C")$amscitperyear),10000,n_c)

sum(dist1 < val1)/length(dist1)
sum(dist1 < val2)/length(dist2)
sum(dist1 < val3)/length(dist3)
```

The induced p-value for hypothesis 1 is 0. The induced p-value for hypothesis 2 is 0.07788. The induced p-value for hypothesis 3 is 0. Hence we reject hypothesis 1 and 3, and conclude that $\mu(A_{citperyear}) < \mu(B_{citperyear}) \leq \mu(C_{citperyear})$.

@@

```{r, include = FALSE, echo = FALSE}
muA <- mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B"))$amscitperyearpower, na.rm = TRUE)
muB <- mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"))$amscitperyearpower, na.rm = TRUE)
muC <- mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C"))$amscitperyearpower, na.rm = TRUE)

@@

set.seed(0)
n_a <- length(na.omit(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyearpower))
n_b <- length(na.omit(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyearpower))
n_c <- length(na.omit(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&institution=="domesticr1"&field=="math"&(role=="professor"))$amscitperyearpower))

pop<- filter(df, ((role=="professor")&field=="math"&institution=="domesticr1"))

dist1 <- meanPermutation(na.omit(filter(pop,(lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "B and C"|lettergroup == "B Only"))$amscitperyearpower),10000,n_a)
dist2 <- meanPermutation(na.omit(filter(pop,lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B"|lettergroup == "C Only")$amscitperyearpower),10000,n_b)
dist3 <- meanPermutation(na.omit(filter(pop,lettergroup == "A Only"|lettergroup == "A and B"|lettergroup == "C Only"|lettergroup == "B and C")$amscitperyearpower),10000,n_c)

sum(dist1 < val1)/length(dist1)
sum(dist2 < val2)/length(dist2)
sum(dist3 < val3)/length(dist3)
```

The induced p-value for hypothesis 1 is 0. The induced p-value for hypothesis 2 is 0.0974. The induced p-value for hypothesis 3 is 0.0002. Hence we fail to reject hypothesis 2 at a 2% significance level and reject hypotheses 1 and 3 in favor of the alternative. We conclude that after adjusting for age $\mu(A) < \mu(B) \leq \mu(C)$.

@@

```{r, echo = FALSE, include = FALSE}
mean(filter(df, (lettergroup == "A Only"|lettergroup == "A and B")&role=="professor")$amscitperyear, na.rm = TRUE)
mean(filter(df, (lettergroup == "B and C"|lettergroup == "B Only"|lettergroup == "A and B")&role=="professor")$amscitperyear, na.rm = TRUE)
mean(filter(df, (lettergroup == "C Only"|lettergroup == "B and C")&role=="professor")$amscitperyear, na.rm = TRUE)
```

@@

2. For letter B - 41.86467

### 2020-01-29 index.tex
revision
update

  \title{Response to Pachter's Review}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Joshua Paik and Igor Rivin}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{1/22/2020}

@@

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\begin{quote}
\emph{``Academic politics are so vicious precisely because the stakes
are so small.''} - Henry Kissinger
\end{quote}

\begin{quote}
\emph{``The value of preprints is in their ability to accelerate
research via the rapid dissemination of methods and discoveries.''} -
\href{https://liorpachter.wordpress.com/2019/10/21/zero-data-rna-seq/}{Lior
Pachter}
\end{quote}

Last Update: January 23rd 2:50 A.M.

@@

We would like to thank Lior for finding the bug in our appendix which
pushes the mean Google Scholar citations of B further away from A. We
agree that the sentence - ``while this is not optimal, a quick sample
size calculation shows that one needs 303 samples or 21\% of the data to
produce statistics at a 95\% confidence level and a 5\% confidence
interval.'' - is ridiculous.

@@

We will compare the mean number of citations and citations per year
between signers of Letters A, B, and C. We will validate the
significance of the difference between signers using a
\href{https://en.wikipedia.org/wiki/Resampling_(statistics)\#Permutation_tests}{permutation
test}.

The following is copy and pasted from our paper (with edit in
parentheses):

A permutation test is a nonparametric way of assessing the difference in
mean between two populations. We are interested in whether an observed
difference in mean is due to chance, and we can assess this in the
following way.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{-1}
\tightlist
\item
  Record the true difference in mean (or median) (\(d\mu\)).
\item
  \(H_0: d\mu = 0, H_1: d\mu < 0\)
\item
  Sample without replacement 1/2 of the combined data set (X) and what
  is left (Y)
\item
  Take the mean (or median) of X and Y and record the difference
\item
  Repeat 10,000 times and plot the histogram
\item
  Record the number of points (m) in the induced distribution that is
  more extreme than or equal to the observed \(d\mu\). The probability
  m/10,000 is the probability that what was observed was due to chance.
\end{enumerate}

@@

In the following figure, the vertical red bar is the observed difference
for hypothesis 1 in the induced distribution, blue is the observed
difference for hypothesis 2, and green is the observed difference for
hypothesis 3.

\includegraphics{index_files/figure-latex/unnamed-chunk-12-1.pdf}

The induced p-value for hypothesis 1 is 0. The induced p-value for
hypothesis 2 is 0.0002. The induced p-value for hypothesis 3 is 0. Hence
we reject all three null hypotheses in favor of the alternative, and
\(\mu(A) < \mu(B) < \mu(C)\).

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-14-1.pdf}

@@

As above, the vertical red bar is the observed difference for hypothesis
1 in the induced distribution, blue is the observed difference for
hypothesis 2, and green is the observed difference for hypothesis 3.

\includegraphics{index_files/figure-latex/unnamed-chunk-16-1.pdf}

The induced p-value for hypothesis 1 is 0. The induced p-value for
hypothesis 2 is 0.0099. The induced p-value for hypothesis 3 is 0. Hence
we reject all three null hypotheses, even assessing the induced p-value
at a 2\% significance level, in favor of the alternative, and we
conclude that
\(\mu(A_{citperyear}) < \mu(B_{citperyear}) < \mu(C_{citperyear})\).

One may object of our usage of the mean here, opposed to the median. So
we reperform the permutation test with the median. All three induced
p-values are 0 so we can reject all three null hypotheses when comparing
median citations per year.

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-22-1.pdf}

@@

We plot the Age vs Citations per Year for all math R1. We generate a
linear regression model and output a 95\% confidence interval for the
slope. \includegraphics{index_files/figure-latex/unnamed-chunk-27-1.pdf}

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-30-1.pdf}

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-32-1.pdf}

@@

As above, the vertical red bar is the observed difference for hypothesis
1 in the induced distribution, blue is the observed difference for
hypothesis 2, and green is the observed difference for hypothesis 3.

\includegraphics{index_files/figure-latex/unnamed-chunk-34-1.pdf}

The induced p-value for hypothesis 1 is 0. The induced p-value for
hypothesis 2 is 0.0207. The induced p-value for hypothesis 3 is 0. Hence
we fail to reject hypothesis 2 at a 2\% significance level and reject
hypotheses 1 and 3 in favor of the alternative. We conclude that after
adjusting for age \(\mu(A) < \mu(B) \leq \mu(C)\).

@@

This method was suggested by a friend as a final check to eliminate any
question that age was the greatest confounder. We want to show that
\(\mu(A) < \mu(B \cup C)\). We will randomly sample a population of 20
from A, called \(X\). For each member \(x \in X\), we will find every
person from B and C that is within a two year age interval from \(x\).
We will randomly sample one, and induce a new population \(Y\). Then we
will compare the means by storing X-Y. We repeat this 1,000 times and
plot a histogram of the induced values. If 0 is within this new
distribution, then maybe there is a chance, a totally slim one after
above, that in fact age is a confounder. If the distribution is
primarily negative, then X \textless{} Y. Otherwise X \textgreater{} Y.
We perform this analysis with both AMS citations and Google Scholar
citations.

\includegraphics{index_files/figure-latex/unnamed-chunk-36-1.pdf}

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-38-1.pdf}

@@

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-42-1.pdf}

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-44-1.pdf}

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-45-1.pdf}

So it is clear that Pachter's analysis was some sort of magic trick,
potentially a thought experiment, and a misleading one. It is highly
unlikely that a tenured and respected expert in computation and
statistics did not know the above result, expecially when a student he
suggests take an introductory statistics course immediately spotted it.
We claim and assert that he purposefully chose his 36 cutoff to try to
undermine our results.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  For letter A - 15.98726
\item
  For letter B - 41.12345
\item
  For letter C - 55.5451
\end{enumerate}

Temple has an average citations per year of 12.33, so we retract our
claim that letter A is comparable to Temple. It is closer to the
University of Massachusetts - Amherst which has an average citations per
year of 16.17. By
\href{https://www.usnews.com/best-graduate-schools/top-science-schools/mathematics-rankings}{US
News}, University of Massachusetts - Amherst's Math Department has a
rank of 55. Rutgers has an average citations per year of 35.01, so we
retract our claim that letter B is comparable to Rutgers. It is closer
to the University of Minesotta which has an average citations per year
of 42.07 and a US News Ranking of 19. For Letter C, we claimed that it
was another tier higher - indeed it is closer to the University of
Chicago, which has an average citations per year of 56.27, ranked 6 by
US News.

An astute observer would notice we are not exactly comparing apples to
apples. Presumably ones first publication could be before one finishes
their PhD. So even with the boost, the order amongst letter signers
stands.

>>>>

  \title{Response to Pachter's Review}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Joshua Paik and Igor Rivin}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{1/22/2020, Last Update: 1/28/2020 6:00 P.M. G.M.T.}

@@

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\begin{quote}
\emph{``The value of preprints is in their ability to accelerate
research via the rapid dissemination of methods and discoveries.''} -
\href{https://liorpachter.wordpress.com/2019/10/21/zero-data-rna-seq/}{Lior
Pachter}
\end{quote}

@@

We would like to thank Pachter for finding the bug in our appendix which
pushes the mean Google Scholar citations of B further away from A. We
agree that the sentence - ``while this is not optimal, a quick sample
size calculation shows that one needs 303 samples or 21\% of the data to
produce statistics at a 95\% confidence level and a 5\% confidence
interval.'' - is ridiculous.

@@

We will compare the mean number of citations and citations per year
(that is years elapsed since completion of PhD) between signers of
Letters A, B, and C. We will validate the significance of the difference
between signers using a
\href{https://en.wikipedia.org/wiki/Resampling_(statistics)\#Permutation_tests}{permutation
test}.

A permutation test is a non-parametric means of assessing the
significance test of a population. Throughout this section, we will be
comparing the mean citations of two populations, X and Y. We will work
under the assumption that our null hypothesis is
\(H_0: \mu(X) = \mu(Y)\), and our alternative is
\(H_1: \mu(X) < \mu(Y)\).

A permutation test works as follows. Let \(X\) and \(Y\) be our relevant
populations, of size \(n_X\) and \(n_Y\). We would like to know whether
we can accept that the observed difference in means was not due to
chance. We record the observed difference in means as
\(\delta = \mu(X) - \mu(Y)\). We then take the union of our two
population, \(Z = X \cup Y\), and randomly partition \(Z\) into two new
sets \(A\) and \(B\), where \(|A| = n_X\) and \(|B| = n_Y\). We store
\(\mu(X)-\mu(Y)\) and induce a distribution \(D\) of potential
differences and repeat the process \(n=10,000\) times. We can induce the
p-value, or the probability that our observed difference was due to
chance by the probability
\(p = |\{d:d\leq \delta, \forall d \in D\}|/n\).

@@

The induced p-value for hypothesis 1 is 0. The induced p-value for
hypothesis 2 is 0.0016. The induced p-value for hypothesis 3 is 0. Hence
we reject all three null hypotheses in favor of the alternative, and
\(\mu(A) < \mu(B) < \mu(C)\).

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-13-1.pdf}

@@

The induced p-value for hypothesis 1 is 0. The induced p-value for
hypothesis 2 is 0.07788. The induced p-value for hypothesis 3 is 0.
Hence we reject hypothesis 1 and 3, and conclude that
\(\mu(A_{citperyear}) < \mu(B_{citperyear}) \leq \mu(C_{citperyear})\).

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-18-1.pdf}

@@

We plot the Citations per Year vs Age for all math R1. We generate a
linear regression model and output a 95\% confidence interval for the
slope. \includegraphics{index_files/figure-latex/unnamed-chunk-23-1.pdf}

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-26-1.pdf}

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-28-1.pdf}

@@

The induced p-value for hypothesis 1 is 0. The induced p-value for
hypothesis 2 is 0.0974. The induced p-value for hypothesis 3 is 0.0002.
Hence we fail to reject hypothesis 2 at a 2\% significance level and
reject hypotheses 1 and 3 in favor of the alternative. We conclude that
after adjusting for age \(\mu(A) < \mu(B) \leq \mu(C)\).

@@

This method was suggested by a friend as a final check to eliminate any
question that age was the greatest confounder. We want to show that
\(\mu(A) < \mu(B \cup C)\). We will randomly sample a population of 20
from A, called \(X\). For each member \(x \in X\), we will find every
person from B and C that is within a four year age interval from \(x\).
We will randomly sample one, and induce a new population \(Y\). Then we
will compare the means by storing X-Y. We repeat this 1,000 times and
plot a histogram of the induced values. If 0 is within this new
distribution, then maybe there is a chance, a totally slim one after
above, that in fact age is a confounder. If the distribution is
primarily negative, then X \textless{} Y. Otherwise X \textgreater{} Y.
We perform this analysis with both AMS citations and Google Scholar
citations.

\includegraphics{index_files/figure-latex/unnamed-chunk-31-1.pdf}

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-33-1.pdf}

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-35-1.pdf}

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-37-1.pdf}

@@

\includegraphics{index_files/figure-latex/unnamed-chunk-39-1.pdf}

@@

So it is clear that Pachter's analysis was some sort of magic trick,
potentially a thought experiment, and a misleading one. It is highly
unlikely that a tenured and respected expert in computation and
statistics did not know the above result, expecially when a student he
suggests take an introductory statistics course immediately spotted it.
One may suspect that he purposefully chose his 36 cutoff to try to
undermine our results.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  For letter A - 15.98726
\item
  For letter B - 41.86467
\item
  For letter C - 55.5451
\end{enumerate}

Temple has an average citations per year of 12.33, so we retract our
claim that letter A is comparable to Temple. It is closer to the
University of Massachusetts - Amherst which has an average citations per
year of 16.17. By
\href{https://www.usnews.com/best-graduate-schools/top-science-schools/mathematics-rankings}{US
News}, University of Massachusetts - Amherst's Math Department has a
rank of 55. Rutgers has an average citations per year of 35.01, so we
retract our claim that letter B is comparable to Rutgers. It is closer
to the University of Minesota which has an average citations per year of
42.07 and a US News Ranking of 19. For Letter C, we claimed that it was
another tier higher - indeed it is closer to the University of Chicago,
which has an average citations per year of 56.27, ranked 6 by US News.

An astute observer would notice we are not exactly comparing apples to
apples. Presumably one's first publication could be before one finishes
their PhD. So even with the boost, the order amongst letter signers
stands.
